---
title: "DATA 607 - Assignment 3 - R Character Manipulation and Date Processing"
author: "Preston Peck"
date: "9/9/2021"
output: html_document
---

```{r}
library(stringr)
library(dplyr)
library(qdapRegex)
library(purrr)
```

******

# Question 1
#### Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

(?i) is case insensitive

```{r}
read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv") %>%
  filter(str_detect(Major, "(?i)data|statistics"))
```

******

# Question 2
#### Write code that transforms the data below into a format like this:
#### c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
question2 <- read.csv("data/question2.csv", header = FALSE)
question2
```

##### String

```{r}
question2String <- question2

indexesRegex <- "([^A-Za-z()\" ])"

beforeNextRowRegex <- "\"\\\\"
afterNextRowRegex <- "\"\""

aroundItemsRegex <- "\\\\"
betweenItemsRegex <- "\" \""

question2String[[1]] <- question2String[[1]] %>% 
  str_replace_all(indexesRegex, "")  %>% 
  str_squish

question2String %>% 
  paste %>%
  str_replace_all(paste(beforeNextRowRegex, "|", aroundItemsRegex, sep = ""), "")  %>%
  str_replace_all(afterNextRowRegex, "\"") %>%
  str_replace_all(betweenItemsRegex, "\", \"")
```
  
##### Vector

```{r}
question2Vector <- question2

question2Vector %>%
  paste %>%
  str_replace_all(indexesRegex, "") %>%
  rm_between(" \"", "\" ", extract = TRUE) %>%
  unlist %>%
  discard(function(x) x == "")
```

******

# Question 3
#### Describe, in words, what these expressions will match:
##### • (.)\1\1
This may have been incorrectly written\
\\1 can be a reserved symbol, so the intention might have been to escape it (\\\1)\
Currently, it captures any letter, then a \\, then a 1, then a \\, then a 1 (length 5)

```{r}
anySlashOneSlashOneRegex <- "(.)\1\1"
words <- c("a\1\1", "b\1\1", "a\1\1a", "something")
str_extract(words, regex(anySlashOneSlashOneRegex, ignore_case = TRUE))
```

###### If rewritten as: (.)\\\\1\\\\1\
It will capture any 3 of the same characters in sequence (form 111, length 3)

```{r}
sameThreeSequentiallyRegex <- "(.)\\1\\1"
words <- c("aaa", "aaaa", "dad", "daaad", "a aaa", "111")
str_extract(words, regex(sameThreeSequentiallyRegex, ignore_case = TRUE))
```



##### • "(.)(.)\\\2\\\1"
It will capture any 2 characters, then those same characters in reverse sequence all surrounded by quotes (form "1221", length 6)

```{r}
anyTwoThenReversedQuotedRegex <- "\"(.)(.)\\2\\1\""
words <- c("aaaa", "aaa", "dad", "daaad", "daad","1221",
           "\"aaaa\"", "\"aaa\"", "\"dad\"", "\"daaad\"", "\"daad\"", "\"1221\"")
str_extract(words, regex(anyTwoThenReversedQuotedRegex, ignore_case = TRUE))
```



##### • (..)\1
This may have been incorrectly written\
\\1 can be a reserved symbol, so the intention might have been to escape it (\\\1)\
Currently, it captures any 2 characters, then a \\, and then a 1 (length 4)

```{r}
anyTwoThenSlashOneRegex <- "(..)\1"
words <- c("ab\1", "ab\1\1", "aa\1\1a", "something")
str_extract(words, regex(anyTwoThenSlashOneRegex, ignore_case = TRUE))
```

###### If rewritten as: (..)\\\\1\
It will capture any 2 characters, then those same characters again in the same sequence (form 1212, length 4)

```{r}
sameAnyTwoTwiceRegex <- "(..)\\1"
words <- c("aaaa", "aa", "daad", "dada", "adaaadaa", "1212")
str_extract(words, regex(sameAnyTwoTwiceRegex, ignore_case = TRUE))
```



##### • "(.).\\\1.\\\1"
It will capture any 2 characters, then another of the first character, then any character, then another of the first character all surrounded by quotes (form "12131", length 7)

```{r}
anyTwoFirstAnyFirstQuotedRegex <- "\"(.).\\1.\\1\""
words <- c("azaxa", "aaaaa", "aaaa", "abcde", "hahah", "12131",
           "\"azaxa\"", "\"aaaaa\"", "\"aaaa\"", "\"abcde\"", "\"hahah\"", "\"12131\"")
str_extract(words, regex(anyTwoFirstAnyFirstQuotedRegex, ignore_case = TRUE))
```

##### • "(.)(.)(.).*\\\3\\\2\\\1"
It will capture 3 or more characters, then the first three characters in reverse sequence all surrounded by quotes (form "123...321", length min 6)

```{r}
anyNThreeTwoOneQuotedRegex <- "\"(.)(.)(.).*\\3\\2\\1\""
words <- c("racecar", "aracecar", "aaaaaaaaa", "hahahahah", "civic", "123...321",
           "\"racecar\"", "\"aracecar\"", "\"aaaaaaaaa\"", "\"hahahahah\"", "\"123...321\"")
str_extract(words, regex(anyNThreeTwoOneQuotedRegex, ignore_case = TRUE))
```

******

# Question 4
#### Construct regular expressions to match words that:
##### • Start and end with the same character.

\\1 is a look behind to group 1\
^ is beginning of the string\
$ is end of the string\
\\b check for word boundary which can detect whitespace between words or beggining or end of string

```{r}
sameStartAndEndCharacterWordRegex <- "^([A-Za-z]).*\\1$"
sameStartAndEndCharacterWordSentenceRegex <- "\\b([A-Za-z])\\S*\\1\\b"
words <- c("dad", "daddy", "mom", "mommy", "dogged", "I  am a  dad")
str_extract(words, regex(sameStartAndEndCharacterWordSentenceRegex, ignore_case = TRUE))
```

##### • Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
\\S looks for all non-whitespace characters, great for parsing individual words in a string

```{r}
repeatedPairWordRegex <- "\\S*([A-Za-z]{2})\\S*\\1\\S*"
repeatedPairLettersRegex <- "([A-Za-z]{2})\\S*\\1"

words <- c("church", "mychurches", "hahaha", "rolypoly", "chachacha", "A rolypoly, haha!")
str_extract(words, regex(repeatedPairWordRegex, ignore_case = TRUE))
str_extract(str_extract(words, regex(repeatedPairLettersRegex, ignore_case = TRUE)), "^.{2}")
```